---
title: "MTWINS Exposure to Community Violence Project"
subtitle: "Interaction Analysis"
author: "Gabriela Suarez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Load packages
```{r load_packages, message=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(interactions)
```

# 2. Load data
```{r data, message=FALSE}
cov = read_csv("mtwins_ecv_covariates.csv")
amy = read_csv("main_effects_scan-cov_extracted_data.csv")
```

## 2.1. Prep & Merge datasets
```{r prep, message=FALSE}
cov[cov==-99] = NA
df = full_join(cov, amy, by="MTWINSID") 
```

## 2.2. Rename Variables
```{r rename, message=FALSE}
df = df %>%
  rename(ADI = ADINAT) %>%
  rename(Violence_Exposure = INDV) %>%
  rename(RT_Amygdala_Threat = RTBL009) %>%
  rename(LF_Amygdala_Threat = LFBL009) %>%
  rename(RT_Amygdala_Neutral = RTBL015) %>%
  rename(LF_Amygdala_Neutral = LFBL015) %>%
  rename(Parental_Nurturance = TRINVOL) %>%
  rename(Family_Income = ANNINC) %>%
  rename(Parent_Education = PCEDUC) %>%
  rename(Twin_Age = TWAGE) %>%
  rename(Twin_Gender = TWSEX) %>%
  rename(Twin_Race = TWRACER)
```

# 3. Parental Nurturance

```{r hist_nurture, warning=FALSE, message=FALSE}
# Calculate mean and standard deviation
mean_val = mean(df$Parental_Nurturance, na.rm = TRUE)
sd_val = sd(df$Parental_Nurturance, na.rm = TRUE)

# Create a histogram with different colors for bars within +/- 1 SD from the mean
hist = df %>%
  ggplot(aes(x = Parental_Nurturance)) + 
    geom_histogram(aes(fill = cut(Parental_Nurturance, 
                                  breaks = c(-Inf, mean_val - sd_val, mean_val + sd_val, Inf),
                                  labels = c('-1 SD', 'Mean', '+1 SD'))),
                   binwidth = 1, colour = 'black') + 
    scale_fill_manual(values = c("#000099", "#56B4E9", "#FFCC00"), name = 'Parental\nNurturance') + 
    labs(x = 'Parental Nurturance', y = 'Count', ) + 
    theme_classic() + theme(text = element_text(size = 14, family = "sans")) + 
  theme(legend.position = "top")
hist

ggsave("nurture_hist.pdf", plot=hist, width = 4.5, height = 4, unit = "in", dpi=1200)
```


# 4. Interactions

## 4.1. a Path: Parental involvement moderates the link between ADI and ECV

Visualize using package "interactions"

```{r a_path_out}
aPath = lm(data=df, Violence_Exposure ~ ADI * Parental_Nurturance)
summary(aPath)
```

### 4.1.1. Johnson-Neyman Plot
```{r a_path_johnson-neyman, warning=FALSE, message=FALSE}
sim_slopes(aPath, pred = ADI, modx = Parental_Nurturance, jnplot = TRUE)

#Run this line of code with the predictor and the moderator reversed to get the region of significance
sim_slopes(aPath, pred = Parental_Nurturance, modx = ADI, jnplot = TRUE) 

interact_plot(aPath, pred = ADI, modx = Parental_Nurturance, 
              interval = TRUE, int.width = 0.9, 
              rug=TRUE, rug.sides = "bl")
```

### 4.1.2. Interaction Plot: Powerpoint Formatting
```{r a_path_interact_ppt, warning=FALSE, message=FALSE}
MIpalette = c("#000099", "#56B4E9", "#FFCC00")

#This code creates a gray box to overlay on the interaction plot to show region of significance
#Use the interval that you found from the simp_slopes code above for xmin and xmax
ros = data.frame(xmin = 50.12,
                 xmax = 99,
                 ymin = -Inf, 
                 ymax = Inf) #This code gets used in the interact_plot code - see geom_rect
p1 = interact_plot(aPath, pred = ADI, modx = Parental_Nurturance, 
                    interval = TRUE, int.width = 0.95, 
                    rug = TRUE, rug.sides = 'bl', 
                    modx.labels = c("-1 SD","Mean","+1 SD"), colors = MIpalette, 
                    x.label = "Area Deprivation Index", 
                    y.label = "Exposure to Community Violence", 
                    legend.main = "Parental Nurturance", data = df) + 
  geom_rect(data = ros, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "lightgray", linetype = 0, alpha = .3, inherit.aes = FALSE) + 
  theme_bw() + theme(text = element_text(size = 18, family = "Fira Sans")) + 
  theme(legend.position = "top")
p1

ggsave("aPath_interact.png", plot=p1, width = 15, height = 15, unit = "cm", dpi=600)
```

### 4.1.3. Interaction Plot: Manuscript Formatting
```{r a_path_interact_manuscript, warning=FALSE, message=FALSE}
MIpalette = c("#000099", "#56B4E9", "#FFCC00")

#This code creates a gray box to overlay on the interaction plot to show region of significance
#Use the interval that you found from the simp_slopes code above for xmin and xmax
ros = data.frame(xmin = 50.12,
                 xmax = 99,
                 ymin = -Inf, 
                 ymax = Inf) #This code gets used in the interact_plot code - see geom_rect

a = interact_plot(aPath, pred = ADI, modx = Parental_Nurturance, 
                    interval = TRUE, int.width = 0.95, 
                    rug = TRUE, rug.sides = 'bl', 
                    modx.labels = c("-1 SD","Mean","+1 SD"), colors = MIpalette, 
                    x.label = "Area Deprivation Index", 
                    y.label = "Exposure to\nCommunity Violence", 
                    legend.main = "Parental\nNurturance", data = df) + 
  geom_rect(data = ros, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "lightgray", linetype = 0, alpha = .3, inherit.aes = FALSE) + 
  theme_classic() + theme(text = element_text(size = 14, family = "sans")) + 
  theme(legend.position = "top")
a

ggsave("aPath_interact.png", plot=a, width = 4.5, height = 4, unit = "in", dpi=600)
```

### 4.1.4. How many participants lived in neighborhoods with ADI scores above 50.12 and also reported parental nurturance scores below 45.43?
```{r a_path_table}
df = df %>%
  mutate(inter = if_else(ADI > 50.12 & Parental_Nurturance < 45.43, 1, 0))
table(df$inter)
prop.table(table(df$inter))
```


## 4.2 b Path: Parental involvement moderates the link betwee ECV and amygdala reactivity

Visualize using package "interactions"

```{r b_path_out}
bPath <- lm(data=df, RT_Amygdala_Threat ~ Violence_Exposure * Parental_Nurturance)
summary(bPath)
```

### 4.2.1. Johnson-Neyman Plot
```{r b_path_johnson-neyman, warning=FALSE, message=FALSE}
sim_slopes(bPath, pred = Violence_Exposure, modx = Parental_Nurturance, jnplot = TRUE)

#Run this line of code with the predictor and the moderator reversed to get the region of significance
sim_slopes(bPath, pred = Parental_Nurturance, modx = Violence_Exposure, jnplot = TRUE) 

interact_plot(bPath, pred = Violence_Exposure, modx = Parental_Nurturance, 
              interval = TRUE, int.width = 0.9, 
              rug=TRUE, rug.sides = "bl")
```

### 4.2.2. Interaction Plot: Powerpoint Formatting
```{r b_path_interact_ppt, warning=FALSE, message=FALSE}
MIpalette = c("#000099", "#56B4E9", "#FFCC00")

#This code creates a gray box to overlay on the interaction plot to show region of significance
#Use the interval that you found from the simp_slopes code above for xmin and xmax
ros = data.frame(xmin = 5.72,
                 xmax = 28,
                 ymin = -Inf, 
                 ymax = Inf) #This code gets used in the interact_plot code - see geom_rect
p1 = interact_plot(bPath, pred = Violence_Exposure, modx = Parental_Nurturance, 
                    interval = TRUE, int.width = 0.95, 
                    rug = TRUE, rug.sides = 'bl', 
                    modx.labels = c("-1 SD","Mean","+1 SD"), colors = MIpalette, 
                    x.label = "Exposure to Community Violence", 
                    y.label = "Right Amygdala Reactivity\n to Threat", 
                    legend.main = "Parental Nurturance", data = df) + 
  geom_rect(data = ros, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "lightgray", linetype = 0, alpha = .3, inherit.aes = FALSE) + 
  theme_bw() + theme(text = element_text(size = 18, family = "Fira Sans")) + 
  theme(legend.position = "top")
p1

ggsave("bPath_interact.png", plot=p1, width = 15, height = 15, unit = "cm", dpi=600)
```

### 4.2.3. Interaction Plot: Manuscript Formatting
```{r b_path_interact_manuscript, warning=FALSE, message=FALSE}
MIpalette = c("#000099", "#56B4E9", "#FFCC00")

#This code creates a gray box to overlay on the interaction plot to show region of significance
#Use the interval that you found from the simp_slopes code above for xmin and xmax
ros = data.frame(xmin = 5.72,
                 xmax = 28,
                 ymin = -Inf, 
                 ymax = Inf) #This code gets used in the interact_plot code - see geom_rect

b = interact_plot(bPath, pred = Violence_Exposure, modx = Parental_Nurturance, 
                    interval = TRUE, int.width = 0.95, 
                    rug = TRUE, rug.sides = 'bl', 
                    modx.labels = c("-1 SD","Mean","+1 SD"), colors = MIpalette, 
                    x.label = "Exposure to Community Violence", 
                    y.label = "Right Amygdala\nReactivity to Threat", 
                    legend.main = "Parental\nNurturance", data = df) + 
  geom_rect(data = ros, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "lightgray", linetype = 0, alpha = .3, inherit.aes = FALSE) + 
  theme_classic() + theme(text = element_text(size = 14, family = "sans")) + 
  theme(legend.position = "top")
b

ggsave("bPath_interact.png", plot=b, width = 4.5, height = 4, unit = "in", dpi=600)
```

### 4.2.4. How many participants reported violence exposure above 5.72 and also reported parental nurturance scores below 38.96?
```{r b_path_table}
df = df %>%
  mutate(inter = if_else(Violence_Exposure > 5.72 & Parental_Nurturance < 38.96, 1, 0))
table(df$inter)
prop.table(table(df$inter))
```


